---
title: "F1 analysis"
author: "Andrew Tait"
date: "08/03/2021"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(forcats)
library(ggplot2)
library(glue)
library(janitor)
library(readr)
library(skimr)
library(snakecase)
library(stringr)
library(tibble)
library(tidyr)

theme_set(ggthemes::theme_few())

setwd("/data")
```

```{r echo=FALSE, results=FALSE, message=FALSE, warning=FALSE}
constructors <- read_csv("f1_data/constructors.csv") %>% clean_names()
drivers <- read_csv("f1_data/drivers.csv") %>% clean_names()
circuits <- read_csv("f1_data/circuits.csv") %>% clean_names()
races <- read_csv("f1_data/races.csv") %>% clean_names()
pit_stops <- read_csv("f1_data/pitStops.csv") %>% clean_names()
results <- read_csv("f1_data/results.csv") %>% clean_names()
```
## Races

```{r echo=FALSE}
skim(races)
```

## Drivers

Which drivers had the most appearances?

```{r echo=FALSE}
drivers %>% 
  inner_join(results, by = "driver_id") %>% 
  mutate(surname = ifelse(code == "RAI", "Räikkönen", surname)) %>% 
  mutate(surname = ifelse(driver_id == 119, "Patrese", surname)) %>% 
  mutate(name = paste(forename, surname)) %>% 
  group_by(driver_id, name) %>% 
  tally(name = "appearances") %>% 
  ungroup() %>% 
  select(name, appearances) %>% 
  arrange(desc(appearances))
```

## Constructors

Which constructors have won the most races?

```{r echo=FALSE}
constructors %>% 
  inner_join(results, by = "constructor_id") %>% 
  filter(position == 1) %>% 
  count(name, name = "wins") %>% 
  arrange(desc(wins))
```

How have constructors' fortunes changed over the years?

```{r echo=FALSE, message=FALSE}
constructors %>% 
  inner_join(results, by = "constructor_id") %>% 
  inner_join(races, by = "race_id") %>% 
  select(name = name.x, year, points) %>% 
  group_by(name, year) %>% 
  summarise(season_points = sum(points)) %>% 
  filter(name %in% c("Ferrari", "McLaren", "Mercedes", "Red Bull", "Williams")) %>% 
  ggplot(aes(x = year, y = season_points)) +
  geom_line(aes(color = name)) +
  labs(title = "Constructors' total points", color = "Constructors") +
  xlab("Season") +
  ylab("Points")
```

## Circuits

Which circuits have hosted the most races?

```{r echo=FALSE}
circuits %>% 
  inner_join(races, by = "circuit_id") %>% 
  select(name = name.x) %>% 
  count(name, name = "races") %>% 
  arrange(desc(races))
```

## Pit stops

How did pit stop times change over the years?

```{r echo=FALSE}
pit_stops %>% 
  inner_join(races, by = "race_id") %>% 
  filter(circuit_id == 9) %>% 
  mutate(duration = milliseconds / 1000) %>% 
  select(year, duration) %>%
  group_by(year) %>% 
  summarise(average_duration = median(duration)) %>% 
  ggplot(aes(x = year, y = average_duration)) +
  geom_line() +
  scale_x_continuous(breaks = 2011:2017, minor_breaks = NULL) +
  labs(title = "Silverstone pit stops") +
  xlab("Season") +
  ylab("Median pit stop (seconds)")
```

Did the consistency of pit stops change over the seasons?

```{r echo=FALSE}
pit_stops %>% 
  inner_join(races, by = "race_id") %>% 
  filter(circuit_id == 9) %>% 
  mutate(year = fct_reorder(as_factor(year), -year), duration = milliseconds / 1000) %>% 
  select(year, duration) %>%
  filter(duration < 50) %>% 
  ggplot(aes(x = year, y = duration)) +
  geom_boxplot() +
  labs(title = "Silverstone pit stops") +
  xlab("Season") +
  ylab("Median pit stop (seconds)") +
  coord_flip()
```

## Economist theme

Economist reader?

```{r echo=FALSE}
pit_stops %>% 
  inner_join(races, by = "race_id") %>% 
  filter(circuit_id == 9) %>% 
  mutate(year = fct_reorder(as_factor(year), -year), duration = milliseconds / 1000) %>% 
  select(year, duration) %>%
  filter(duration < 50) %>% 
  ggplot(aes(x = year, y = duration)) +
  geom_boxplot() +
  labs(title = "Silverstone pit stops") +
  xlab("Season") +
  ylab("Median pit stop (seconds)") +
  coord_flip() +
  ggthemes::theme_economist()
```

## WSJ theme

WSJ reader?

```{r echo=FALSE}
pit_stops %>% 
  inner_join(races, by = "race_id") %>% 
  filter(circuit_id == 9) %>% 
  mutate(year = fct_reorder(as_factor(year), -year), duration = milliseconds / 1000) %>% 
  select(year, duration) %>%
  filter(duration < 50) %>% 
  ggplot(aes(x = year, y = duration)) +
  geom_boxplot() +
  labs(title = "Silverstone pit stops") +
  xlab("Season") +
  ylab("Median pit stop (seconds)") +
  coord_flip() +
  ggthemes::theme_wsj()
```
